---
title: "Report"
output: html_document

---


```{r include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      fig.width = 6,
                      fig.asp = 0.8,
                      out.width = "80%"
                      )
library(tidyverse)
library(forcats)
library(kableExtra)
library(janitor)
```

```{r include=FALSE}

source("scripts/eyespots_exp1.R")

```

## Results:

The ultimate question addressed in this report is does the eyespot ‘sparkle’ decrease the predation rate of lepidoptera. For each experiment, I used a binomial logit-link generalized linear model, with quasi-likelihoods to account for overdispersion, to analyse the effect of design, location, time (collection) and temperature on the likelihood of predation. An interaction term between location and collection was included in the model for experiment 1. All analyses and data cleaning were carried out in R (ver 4.3.3) with the tidyverse range of packages (Wickham et al, 2019), car (Fox et al, 2023), patchwork (Lin et al, 2024) and ggeffects (Lüdecke et al, 2024) packages. Model residuals were checked with the performance package (Lüdecke et al, 2021). Summary tables were produced with broom (Robinson et al 2022) and KableExtra (Zhu, 2020) packages. In the analysis, the term collection refers to the event of collecting in the models to count predation levels. This happened 9 times throughout each experiment. Collection is used as a reference for time into the experiment in analyses and is used as a numerical variable, so that any potential effects of learning can be investigated. 


# Influence of Design on Predation:

## Experiment 1:

The main hypothesis of this experiment was that design 3 (2 eyespots) would show the least amount of predation compared to designs 1 (blank) and 2 (1 eyespot), with blank models showing the greatest likelihood of predation. Previous research has shown that blank models show a significantly greater probability of predation compared to spotted models (Stevens et al, 2007) . 

```{r echo=FALSE, fig.asp=0.8, fig.pos="H", fig.width=7, warning=FALSE, message=FALSE}

exp1_model %>% kable_styling(latex_options = "HOLD_position")

```
**Table 1** Eyespot models show significantly lower log odds likelihood of predation compared to blank models. Output displaying estimated coefficients with 95% confidence intervals, z-values, and associated p-values from a binomial logit-link generalized linear model assessing the predictors of predation. The predictors include collection, location (2 levels), design (3 levels), temperature and an interaction between collection and location.  

There was a significant overall effect of design on likelihood of predation (χ² 2,1050 = 67.96, P = <0.0001), with blank models showing a mean predation probability of 84.23% [95% CI: 80.04 - 87.68%]. Mean probability of predation for models with 1 eyespot was 62.57% [95% CI: 56.95-67.87%], and the lowest mean probability of predation was 58.43% for models with 2 eyespots [95% CI: 52.75-63.89%].

Post hoc analysis showed that these differences were statistically significant for blank models compared to 1 eyespot models (Tukey test:zInf = 6.43, P = <0.0001) and 2 eyespot models (zInf = 7.41, P = <0.0001). The analysis showed that the odds of predation for blank models was 3.2 times higher than design 2 (SE = 0.58, z = 6.43, p <0.0001) and 3.8 times higher than design 3 (SE = 0.69, z = 7.41, p <0.0001). There was no overall evidence of a difference between 1 eyespot models and 2 eyespot models (SE = 0.19, z = 1.08, p = 0.529). 

```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

locdes_plot

```
**Figure 1:** Probability of predation higher for all designs at location 2. Plot shows probabilities of predation across three designs of artificial Lepidoptera. Pattern of predation across all three designs remain similar across both locations but a greater probability of predation seen at location 2. Upper and lower boundaries for each data point represent 95% confidence intervals. Key describes which colours represent each design. Y axis represents probability of predation and x axis represents location. Predictions made using the ggpredict function which analysed a binomial logit-link generalized linear model.

## Experiment 2:

The main hypothesis for this experiment was that design 1 ('sparkle' within eyespots) would show the lowest likelihood of predation, and that design 3 (control - 2 eyespots, no 'sparkle') would show the highest likelihood of predation compared to design 1 and design 2 ('sparkle' outside of eyespots). Previous research concluded that models with eyespots containing a 'sparkle' have a lower predation rate compared to eyespots without (Blut et al, 2012). It was thought that models with 'sparkles' outside of the eyespots would show lower predation rate compared to the control because of an extra level of conspicuousness. 

```{r echo=FALSE, fig.asp=0.8, fig.pos="H", fig.width=7, warning=FALSE, message=FALSE}

exp2_model %>% kable_styling(latex_options = "HOLD_position")

```
**Table 2** No statistically significant evidence of a difference in effect of model designs on predation. Output displaying estimated coefficients with 95% confidence intervals, z-values, and associated p-values from a binomial logit-link generalized linear model assessing the predictors of predation. The predictors include collection, location (2 levels), design (3 levels) and temperature.

Design did not have a significant effect on the likelihood of predation (χ² 2,1033 = 3.1, P = 0.2). A descriptive summary of the raw data indicated there may be a difference in predation levels of models containing a 'sparkle' compared to other designs at location 1, but in location 2, predation levels were almost equal for all designs. Due to this observation, an interaction term between location and design was added to the model. The indication of a potential interaction was supported by the predicted probabilities of predation in each location (figure 2).

```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

locdes3_plot

```
**Figure 2** Models with eyespots containing 'sparkle' predicted less likely to be predated, but only in location 1. Despite the visual difference, it is not statistically significant. Plot shows the probability of predation for three designs of artificial lepidoptera at two locations. Models were left exposed for 24 hours. Y axis represents predation probability and x axis represents design type. Probability predictions made using the gg_predict function which analysed a binomial logit-link generalized linear model. Analysis of the model did not show a significant effect of design on probability of predation. 

Adding the interaction term enabled pairwise comparisons between design to be made within each location. As expected, post hoc analysis did not show a significant difference in the odds of predation between designs at location 2. In location 1, comparisons between designs did indicate that predation was approximately 0.65 times less likely to occur in models with eyespots containing 'sparkles'. However, these differences were not large enough to be significant (design 1 - design 2: SE = 0.15, z = -1.87, P = 0.15)(design 1 - design 3: SE = 0.14, z = -1.99, P = 0.11). This is supported by the interaction between location and design not having a significant effect on predation (χ² 2,1033 = 2.01, P = 0.37). 
 
# Experiment 3:

The main hypothesis of this experiment was that design 3 (control - 2 eyespots without a 'sparkle') would show the highest likelihood of predation compared to the two 'sparkle' designs. It was hypothesised that design 1 (UV-reflective 'sparkle') would show a lower likelihood of predation compared to design 2 (UV-absorbant 'sparkle') because 'sparkles' in lepidoptera are UV-reflective (Blut et al, 2012). 

```{r echo=FALSE, fig.asp=0.8, fig.pos="H", fig.width=7, warning=FALSE, message=FALSE}

exp3_model %>% kable_styling(latex_options = "HOLD_position")

```
**Table 2** Control model shows significantly higher likelihood of predation compared to model with UV-reflective 'sparkles' within eyespots. Output displaying estimated coefficients with 95% confidence intervals, z-values, and associated p-values from a binomial logit-link generalized linear model assessing the predictors of predation. The predictors include collection, design (3 levels) and temperature.

There was a significant overall effect of design on likelihood of predation (χ² 2,1051 = 11.40, P = <0.001), with UV-reflective models showing a mean predation probability of 52.7% [95% CI: 47.3 - 58.1%]. Mean probability of predation for UV-absorbant models was 49.2% [95% CI: 43.8-54.6%], and the highest mean probability of predation was 61.8% for control models [95% CI: 56.5-66.9%].

Post hoc analysis showed that these differences were statistically significant for control models compared to UV-reflective models (Tukey test:zInf = -2.36, P = <0.05) and UV-absorbant models (zInf = -3.25, P = <0.01). The analysis showed that the odds of predation for UV-reflective models was 0.69 times lower than control models (SE = 0.11, z = -2.36, p <0.05) and the odds of predation for UV-absorbant models was 0.60 times lower than control models (SE = 0.09 z = -3.25, p <0.01). There was no overall evidence of a difference between both 'sparkle' models (SE = 0.18, z = 0.90, p = 0.64). 

```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

exp3des_plot

```


## Influence of location on Predation

Prior to experiment 1, it was hypothesised that the level of predation between the two locations would follow the same pattern because predation levels were similar during the preliminary experiment. Analysis of the model for experiment 1 showed that likelihood of predation was significantly lower in location 2 compared to location 1 (logit-odds: -1.76 [95% CI: -2.44 - -1.09], z = -5.12, df = 1050, P = <0.0001). Predicted probabilities of predation for models in location 1 was

## Influence of Learning on Predation: 

Previous studies have found that over time birds are capable of learning that the lepidopteran models are fake prey (Blut et al, 2012). It was hypothesised that the level of predation would increase over the experimental period because of this. It was plausible that the effects of learning could be different between the two locations but also between designs, therefore two interaction terms, one between collection and location and the other between collection and design, were included. I found no significant evidence of an interaction between collection and design (χ² 2,1050 = 3.19, P = 0.2) and this was dropped from the model. I did find significant evidence of an interaction between collection and location (χ² 1,1050 = 54.81, P = <0.001). I found a significant positive relationship between collection and probability of predation at location 2 (logit-odds = 0.48 [95% CI: 0.35 - 0.62], z = 6.86, d.f = 1050, P = <0.0001). In contrast, at location 1, the effect of collection on predation was not statistically significant. This suggests that learning occurred in birds at location 2 but not location 1, which could explain the higher level of predation at location 2.

```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

learning_plot

```
**Figure 4:** The impact of learning on predation. The plot shows the relationship between probability of predation and time into the experiment. Y axis represents probability of predation and x axis represents collection event number. Blue represents design 1 (blank), red represents design 2 (1 eyespot) and green represents design 3 (2 eyespots). Artificial Lepidopteran models were exposed to predators in 48 hour periods across 9 collection events. At location 2, predation levels were significantly influenced by collection (or time into the experiment). This indicates that learning significantly influenced the probability of predation at location 2 but not location 1. Despite this, there is an increase in probability of predation after collection 2 at location 1. The shaded zones represent the 95% confidence intervals of predictions for each design. Predictions were made using the augment_glm function which analysed a binomial logit-link generalized linear model.

In experiment 2 ...

## Influence of Temperature on Predation:

The relationship between temperature and predation rates in birds is complex and can be influenced by a combination of factors including metabolic rate, foraging efficiency, breeding behavior, prey behavior, and habitat availability (Alatalo, R.V., 1982). In experiment 1 model analysis showed a significant negative relationship between temperature and probability of predation (logit-odds = -0.21 [95% CI: -0.32 - -0.10], z = -3.85, df = 1050, P = <0.0001). Figure 2 shows the relationship between temperature and predicted probability of predation in experiment 1.


```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

temp

```
**Figure 2** Probability of predation decreases with increases in temperature for experiment 1. Probability of predation decreases from 84% (95% CI: 78 - 88%) at 9 degrees celsius to 69% (95% CI: 60 - 76%) at 13 degrees celsius. Black line shows this decrease and grey area represents 95% confidence intervals of predictions. Y axis represents the predicted probabilities of predation, obtained using the ggpredict function, which analysed a binomial logit-link generalized linear model.

In experiment 2 ...