---
title: "Report"
output:
  pdf_document: default
  html_document: default
---


```{r include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.align = "center",
                      fig.width = 6,
                      fig.asp = 0.8,
                      out.width = "80%"
                      )
library(tidyverse)
library(forcats)
library(kableExtra)
library(janitor)
```

```{r include=FALSE}

source("scripts/eyespots.R")

```

## Results:

The ultimate question addressed in this experiment is does the eyespot ‘sparkle’ decrease the predation rate of lepidoptera. For each experiment, I used a binomial logit-link generalized linear model to analyse the effect of design, location, time (collection) and temperature on the likelihood of predation. All analyses and data cleaning were carried out in R (ver 4.3.3) with the tidyverse range of packages (Wickham et al, 2019), car (Fox et al, 2023), patchwork (Lin et al, 2024) and ggeffects (Lüdecke et al, 2024) packages. Model residuals were checked with the performance package (Lüdecke et al, 2021). Summary tables were produced with broom (Robinson et al 2022) and KableExtra (Zhu, 2020) packages. In the analysis, the term collection refers to the event of collecting in the models to count predation levels. This happened 9 times throughout each experiment. Collection is used as a reference for time into the experiment in analyses and is used as a numerical variable, so that any potential effects of learning can be investigated. 

# Experiment 1:

## Influence of Design on Predation:

The main hypothesis of this experiment was that design 3 (2 eyespots) would show the least amount of predation and design 1 (blank) would show the greatest likelihood of predation because this pattern is seen in previous research (). Compared to the blank design, both 1 eyespot (logit-odds = -1.162 [95% CI: -1.52 - -0.810], z = -6.408, d.f = 1050, P = <0.001) and 2 eyespot (logit-odds = -1.335 [95% CI: -1.69 - -0.985], z = -7.381, d.f = 1050, P = <0.001) designs showed significantly lower log odds likelihood of predation. The size of these differences is clearer when visualised through probabilities (Table 1). Blank models showed a mean probability of predation of 84.23% [95% CI: 80.04 - 87.68%]. Mean probability of predation for models with 1 eyespot was 62.57% [95% CI: 56.95-67.87%], and the lowest mean probability of predation was 58.43% for models with 2 eyespots [95% CI: 52.75-63.89%]. 

```{r echo=FALSE, fig.asp=0.8, fig.pos="H", fig.width=7, warning=FALSE, message=FALSE}

design_table %>% kable_styling(latex_options = "HOLD_position")

```
**Table 1:** Blank models predated significantly more than eyespot models. Design 1 (blank), design 2 (1 eyespot), design 3 (2 eyespots). Values obtained using emmeans function, which analysed a binomial logit-link generalized linear model, to compare the mean probabilities of predation across three models types of artificial lepidoptera. The SE column represents standard error, and lower and upper 95% confidence intervals are shown.  

```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

combined_design_plot

```
**Figure 1:** Blank models show greater probability of predation throughout experiment. Plot shows the probability of predation for three designs of artificial lepidoptera (blank = blue, 1 eyespot = red, 2 eyespots = green) throughout the experiment. Models were left exposed to predators in 48 hour periods. Y axis represents the probability of predation and x axis represents collection (or time into experiment). Shaded areas reprsent the 95% confidence intervals for probability predictions of predation for each design. Probability predictions made using the augment_glm function which analysed a binomial logit-link generalized linear model.

## Influence of Temperature on Predation:

The relationship between temperature and predation rates in birds is complex and can be influenced by a combination of factors including metabolic rate, foraging efficiency, breeding behavior, prey behavior, and habitat availability (Alatalo, R.V., 1982). Model analysis showed a significant negative relationship between temperature and probability of predation (logit-odds = -0.210 [95% CI: -0.318 - -0.103], z = -3.84, df = 1050, P = <0.001). Figure 1 shows the relationship between temperature and predicted probability of predation.


```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

temp_plot

```
**Figure 2** Probability of predation decreases with increases in temperature. Probability of predation decreases from 84% (95% CI: 78 - 88%) at 9 degrees celsius to 69% (95% CI: 60 - 76%) at 13 degrees celsius. Black line shows this decrease and grey area represents 95% confidence intervals of predictions. Y axis represents the predicted probabilities of predation, obtained using the ggpredict function, which analysed a binomial logit-link generalized linear model.

## Influence of location on Predation

It was hypothesised that, over the whole experiment, the level of predation between the two locations would follow the same pattern because predation levels were similar during the preliminary experiment. Analysis of the model showed that likelihood of predation was significantly lower in location 2 compared to location 1 (logit-odds: -1.76 [95% CI: -2.44 - -1.09], z = -5.10, df = 1050, P = <0.001).

```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

locdes_plot

```
**Figure 3:** Probability of predation higher for all designs at location 2. Plot shows probabilities of predation across three designs of artificial Lepidoptera. Pattern of predation across all three designs remain similar across both locations but a greater probability of predation seen at location 2. Upper and lower boundaries for each data point represent 95% confidence intervals. Key describes which colours represent each design. Y axis represents probability of predation and x axis represents location. Predictions made using the ggpredict function which analysed a binomial logit-link generalized linear model.

## Influence of Learning on Predation: 

Previous studies have found that over time birds are capable of learning that the lepidopteran models are fake prey (Blut et al, 2012). It was hypothesised that the level of predation would increase over the experimental period because of this. It was plausible that the effects of learning could be different between the two locations but also between designs, therefore two interaction terms, one between collection and location and the other between collection and design, were included. I found no significant evidence of an interaction between collection and design (χ² 2,1050 = 3.19, P = 0.2) and this was dropped from the model. I did find significant evidence of an interaction between collection and location (χ² 1,1050 = 54.41, P = <0.001). I found a significant positive relationship between collection and probability of predation at location 2 (logit-odds = 0.480 [95% CI: 0.346 - 0.622], z = 6.831, d.f = 1050, P = <0.001). In contrast, at location 1, the effect of collection on predation was not statistically significant. This suggests that learning occurred in birds at location 2 but not location 1.

```{r fig.asp = 0.8, fig.width = 10, echo=FALSE, warning=FALSE, message = FALSE}

learning_plot

```
**Figure 4:** The impact of learning on predation. The plot shows the relationship between probability of predation and time into the experiment. Y axis represents probability of predation and x axis represents collection event number. Blue represents design 1 (blank), red represents design 2 (1 eyespot) and green represents design 3 (2 eyespots). Artificial Lepidopteran models were exposed to predators in 48 hour periods across 9 collection events. At location 2, predation levels were significantly influenced by collection (or time into the experiment). This indicates that learning significantly influenced the probability of predation at location 2 but not location 1. Despite this, there is an increase in probability of predation after collection 2 at location 1. The shaded zones represent the 95% confidence intervals of predictions for each design. Predictions were made using the augment_glm function which analysed a binomial logit-link generalized linear model.
